---
title: "Island Planting - Cleaning and Compiling Data"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(grid)
library(lme4)
```



# Creating a single **GUILD LEVEL** dataframe (2012-2018)

This R Markdown document details the process of cleaning the 2012-2018 Island Planting Project data.  

To begin, we will import all of the .csv files as objects.  
After the data is imported, we clean each year and the compile them together into a single document.  

```{r Import guild-level, include = FALSE}
setwd("C:/Users/zeven_000/Desktop/UCSC/2019 - (3) Summer/Island Planting Exp/Old Data")

D2018 <- read.csv("2018_VegCover.csv")
D2017 <- read.csv("2017_VegCover.csv") 
D2016 <- read.csv("2016_VegCover.csv")
D2015 <- read.csv("2015_VegCover.csv")
D2014 <- read.csv("2014_VegCover.csv")
D2013 <- read.csv("2013_BackgroundVeg.csv")
D2012 <- read.csv("2012_BackgroundVeg.csv")

```


```{r Guild-level restructure and compile, include = FALSE, cache=TRUE, dependson="Import guild-level"}
# ---- 2018 ----
D2018 <- D2018 %>%
  replace(is.na(.), 0) %>%
  mutate(Data.Collector = "JosieJoia",
         Year = "2018",
         nat.grass = Bro.car + Hor.bra + Sti.pul,
         nat.forb = Gri.str + Ach.mil + Sym.chi + Jun.pat) %>%
  rename(Plot = Plot.ID,
         total.cover = Total.cover) %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover)
D2018$Year <- as.factor(D2018$Year)
D2018$Plot <- as.factor(D2018$Plot)


# ---- 2017 ----
D2017 <- D2017 %>%
  mutate(total.cover = (nat.forb + nat.grass + Exo.forb + Exo.grass + Bare + Thatch + Other)) %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover)
D2017$QuadLocation <- recode(D2017$QuadLocation, In = "Island")
D2017$Year <- as.factor(D2017$Year)
D2017$Plot <- as.factor(D2017$Plot)


# ---- 2016 ---- 
D2016 <- D2016 %>%
  replace(is.na(.), 0) %>%
  mutate(Data.Collector = "Rusk",
         Year = "2016",
         total.cover = (Exo.grass + Exo.forb + Thatch + Bare + Brocar + Horbra + Stipul + Gristr + Achmil + Symchi + Junpat),
         nat.grass = Brocar + Horbra + Stipul,
         nat.forb = Gristr + Achmil + Symchi + Junpat) %>%
  mutate(Category = ifelse(Treatment == "Full Mulched", "Full Mulched", ifelse 
                           (Treatment == "Full Not Mulched", "Full Not Mulched", ifelse
                             (Treatment == "Island Mulched", "Island Mulched", "WRONG")))) %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover)
D2016$Plot <- as.factor(D2016$Plot)


# ---- 2015 ----
D2015 <- D2015 %>%
  replace(is.na(.), 0) %>%
  rename(Data.Collector = Data.Collectors,
         Plot = plot..,
         Category = Treatment) %>%
  mutate(total.cover = (Exo.grass + Exo.forb + Thatch + Bare + Bro.car + Hor.bra + Sti.pul + Gri.str + Ach.mil + Sym.chi + Jun.pat),
         nat.grass = Bro.car + Hor.bra + Sti.pul,
         nat.forb = Gri.str + Ach.mil + Sym.chi + Jun.pat,
         Mulch = ifelse(Category == "Full Mulched", "Mulch", ifelse 
                           (Category == "Full Not Mulched", "No Mulch", ifelse
                             (Category == "Island Mulched", "Mulch", "WRONG")))) %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover)
D2015$Plot <- as.factor(D2015$Plot)
D2015$Year <- as.factor(D2015$Year)

  
# ---- 2014 ----
## 2014 has a lot of issues - no plots 9, 14, 15, 19, or 25.

D2014 <- D2014 %>%
  mutate(Planted.guild = ifelse(Quadrant == "NE", "Grass", ifelse 
                                (Quadrant == "NW", "Grass", ifelse
                                  (Quadrant == "SE", "Forb", ifelse
                                    (Quadrant == "SW", "Forb", "WRONG")))),
    total.cover = (nat.forb + nat.grass + Exo.forb + Exo.grass + Bare + Mulch.1 + Gopher + Thatch),
    QuadLocation = "Not recorded") %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover)
D2014$Year <- as.factor(D2014$Year)
D2014$Plot <- as.factor(D2014$Plot)

# ---- 2013 ----
## 2013 has some unique issues - plots 1 and 14 are mislabeled, and need to be fixed. We'll isolate their data, and then group them back in with the correct info. 
D2013_14 <- D2013 %>%
  filter(Plot == "14") %>%
  rename(ID2 = ID,
         Treatment2 = Treatment) %>%
  mutate(ID = recode_factor(ID2, "Island Not Mulched" = "Full Not Mulched"),
            Treatment = recode_factor(Treatment2, "Island" = "Full")) %>%
  select(-ID2, -Treatment2)

D2013 <- D2013 %>%
  filter(Plot != "14") %>%
  full_join(D2013_14)

D2013 <- D2013 %>%
  mutate(Year = "2013") %>%
  filter(Treatment != "Seeded",
         Plot != "13") %>%
  mutate(Category = ifelse(ID == "Island Not Mulched", "Island Not Mulched", 
                           ifelse (ID == "Full Not Mulched", "Full Not Mulched", 
                                   ifelse(ID == "Island Mulched", "Island Mulched",
                                          ifelse(ID == "Full Mulched", "Full Mulched", "WRONG")))),
         Planting.Type = ifelse(Treatment == "Full", "Full",
                                ifelse(Treatment == "Island", "Island", "WRONG")),
         QuadLocation = ifelse(Planting.Type == "Full", "Full", 
                               ifelse(Planting.Type == "Island", "Not recorded", "WRONG")),
         Mow = ifelse(Mow == "Mow", "Mowed",
                      ifelse(Mow == "No Mow", "Unmowed", "WRONG"))) %>%
rename(
    Planted.guild = Grass.or.Forb,
    Mowing.treatment = Mow,
    Exo.grass = Exotic.Grass,
    Exo.forb = Exotic.Forb,
    nat.grass = Native.Grass,
    nat.forb = Native.Forb) %>%
  mutate(total.cover = (nat.forb + nat.grass + Exo.forb + Exo.grass + Bare.Ground.Cover + Mulch.Cover)) %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover) %>%
  filter(Category != "Island Not Mulched")
D2013$Year <- as.factor(D2013$Year)
D2013$Plot <- as.factor(D2013$Plot)


# ---- 2012 ----
D2012 <- D2012 %>%
  replace(is.na(.), 0) %>%
  mutate(total.cover = (nat.forb + nat.grass + Exo.forb + Exo.grass + Bare + Mulch.Cover + Thatch),
         Data.Collector = "Heaston",
         Year = "2012",
         QuadLocation = "Not recorded") %>%
  filter(Treatment != "Seeded",
         Category != "Island No Mulch") %>%
  select(Data.Collector, Year, Plot, Mulch, Planted.guild, Category, QuadLocation, Mowing.treatment, 
         Exo.grass, Exo.forb, nat.grass, nat.forb, total.cover)
D2012$Plot <- as.factor(D2012$Plot)


# Smoosh data together --------------------------------
guilds_12_18 <- full_join(D2018, D2017, by = c("Year", "Data.Collector", "Plot", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                 "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", "total.cover", "QuadLocation"))
guilds_12_18 <- full_join(guilds_12_18, D2016, by = c("Year", "Data.Collector", "Plot", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                  "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", "total.cover", "QuadLocation"))
guilds_12_18 <- full_join(guilds_12_18, D2015, by = c("Year", "Data.Collector", "Plot", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                            "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", "total.cover", "QuadLocation"))
guilds_12_18 <- full_join(guilds_12_18, D2014, by = c("Year", "Data.Collector", "Plot", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                            "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", "total.cover", "QuadLocation"))
guilds_12_18 <- full_join(guilds_12_18, D2013, by = c("Year", "Data.Collector", "Plot", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                            "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", "total.cover", "QuadLocation"))
guilds_12_18 <- full_join(guilds_12_18, D2012, by = c("Year", "Data.Collector", "Plot", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                            "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", "total.cover", "QuadLocation"))

```

After we have a compiled dataset, there are still a few minor cleanup things that need to happen: renaming all of the catergories to be consistent, and renaming the plot IDs to be consistent. 

```{r Guild-level: remove error data, include = FALSE}
guilds_12_18 <- guilds_12_18 %>%
  filter(total.cover > 2.5) %>%
  replace(is.na(.), 0) %>%
  mutate(r.exo.grass = (Exo.grass/total.cover)*100,
         r.exo.forb = (Exo.forb/total.cover)*100,
         r.nat.grass = (nat.grass/total.cover)*100,
         r.nat.forb = (nat.forb/total.cover)*100,
         Category = recode(Category, "Full Mulch" = "Full Mulched",
                           "Full No Mulch" = "Full Not Mulched",
                           "Island Mulch" = "Island Mulched",
                           "Island Mulch " = "Island Mulched"),
         Plot = recode(Plot, "1" = "1 CIM",
                       "12" = "12 CFM",
                       "14" = "14 CFNM",
                       "15" = "15 CIM",
                       "17" = "17 CFNM",
                       "18" = "18 CIM",
                       "19" = "19 CIM",
                       "2" = "2 CFM",
                       "22" = "22 CFM",
                       "24" = "24 CFNM",
                       "25" = "25 CFNM",
                       "4" = "4 CIM",
                       "6" = "6 CFNM",
                       "7" = "7 CFM",
                       "9" = "9 CFM"),
         Mowing.treatment = recode(Mowing.treatment, "Mowed " = "Mowed"),
         QuadLocation = recode(QuadLocation, " " = "Full",
                               "N/A" = "Not recorded"))

```


## Creating a single **SPECIES LEVEL** dataframe (2014-2018)

Species level data was only taken in some years, so we will also produce code to generate species-level dataframes. First, we'll import the data, and then we'll compile it into one dataframe. 



```{r Import species-level, include = FALSE}
setwd("C:/Users/zeven_000/Desktop/UCSC/2019 - (2) Spring/Island Planting Exp/Old Data")
pplots <- read.csv("PlantedPlots.csv")
setwd("C:/Users/zeven_000/Desktop/UCSC/2019 - (3) Summer/Island Planting Exp/Old Data")
D2014 <- read.csv("2014_VegCover.csv")
D2017 <- read.csv("2017_VegCover.csv")

```

```{r Species level restructure and compile, include = FALSE}
pplots <- select(pplots, Year, Data.Collector, Plot, ID, Mulch, Planted.guild, Category, Mowing.treatment, 
                 Planting.Type, QuadLocation, Exo.grass, Exo.forb, nat.grass, nat.forb, 
                 Bro.car, Sti.pul, Hor.bra, Gri.str, Ach.mil, Sym.chi, Jun.pat)
pplots$Year <- as.factor(pplots$Year)
pplots$Category <- as.factor(pplots$Category)
pplots$Mowing.treatment <- as.factor(pplots$Mowing.treatment)
pplots$Planting.Type <- as.factor(pplots$Planting.Type)
pplots$QuadLocation <- as.factor(pplots$QuadLocation)
pplots$Year <- as.factor(pplots$Year)


D2014 <- select(D2014, Data.Collector, Year, Plot, ID, Mulch, Planted.guild, Category, Mowing.treatment, 
                Planting.Type, Exo.grass, Exo.forb, nat.grass, nat.forb, 
                Bro.car, Sti.pul, Hor.bra, Gri.str, Ach.mil, Sym.chi, Jun.pat)
D2014$Year <- as.factor(D2014$Year)

D2017 <- select(D2017, Data.Collector, Year, Plot, ID, Mulch, Planted.guild, Category, Mowing.treatment, 
                Planting.Type, Exo.grass, Exo.forb, nat.grass, nat.forb, 
                Bro.car, Sti.pul, Hor.bra, Gri.str, Ach.mil, Sym.chi, Jun.pat)
D2017$Year <- as.factor(D2017$Year)


species_14_18 <- full_join(pplots, D2014, by = c("Year", "Data.Collector", "Plot", "ID", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                 "Planting.Type", "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", 
                                           "Bro.car", "Sti.pul", "Hor.bra", "Gri.str", "Ach.mil", "Sym.chi", "Jun.pat"))
species_14_18 <- full_join(species_14_18, D2017, by = c("Year", "Data.Collector", "Plot", "ID", "Mulch", "Planted.guild", "Category", "Mowing.treatment", 
                                                        "Planting.Type", "Exo.grass", "Exo.forb", "nat.grass", "nat.forb", 
                                                        "Bro.car", "Sti.pul", "Hor.bra", "Gri.str", "Ach.mil", "Sym.chi", "Jun.pat"))
species_14_18 <- species_14_18 %>%
  replace(is.na(.), 0) %>%
  mutate(Category = recode(Category, "Full Mulch" = "Full Mulched",
                           "Full No Mulch" = "Full Not Mulched",
                           "Island Mulch" = "Island Mulched",
                           "Island Mulch " = "Island Mulched"),
         Mowing.treatment = recode(Mowing.treatment, "Mowed " = "Mowed"),
         QuadLocation = recode(QuadLocation, " " = "Full",
                               "N/A" = "Not recorded"))


```

  
  
  
## Uploading .csv files to GitHub
We will upload our cleaned data to GitHub to be called up easily. 

```{r Upload to GitHub}
write.csv(species_14_18, "Species_2014_2018.csv")
write.csv(guilds_12_18, "Guild_2012_2018.csv")

```
